Community-Conditioned Language Models

### 2020-11-23 paper/paper.tex
spell check

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up we aim at testing the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

@@

All experiments use models with \num{3} sequence encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of paramters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions. 
We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).

@@

\begin{figure*}
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them using using a negative-sampling optimization
algorithm and the message-community co-occurence matrix as ground
truth, using data which spans January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Mean community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm-lmcc-ppl}
\end{figure*}

@@

With indiscernibility values evenly ranging over the whole possible
range (0 to 1), we find that CCLMs can be used to recognize certain
communities based on latent bayesian classification
\ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
distinctive linguistic patterns, none of our model is able to discern
them. In contrast, the distinctive patterns of narrow-topic
communities are just as recognizable by all our LSTM and Transformer
models.

@@

We find that the layer at which the community embedding is taken into
account in a CCLM influences its perplexity \ref{hyp:layer-effect},
albeit weakly. For LSTM models, the perplexity per word, averaged over
all communities, varies between 76.80 and 78.64. (with 80.70 for the unconditioned model)
The perplexity per word is slightly higher for $l_c=0$, and has very
little difference between the other models, with $l_c=1$ being the best.
For transformer models, it varies
between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
function of $l_c$: the maximum is reached at $l_c=2$. We did not
expect non-mononicity, and leave studying this effect to future work.

The pattern of information gain by community is similar across architectures.
Communities that bennifit most from the conditioned model behave that way for 
both the LSTM and transformer.
However, there are some differences.
For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
the best CCLM LSTM but only \num{0.998} for the best transformer model.
Why was the LSTM able to take advantage of the community information for
this community, where the transformer wasn't?
It would be interesting to investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

The representation of communities built by CCLM exhibits a clear
correlation with user-coocurrence patterns
\ref{hyp:extra-linguistic-correlation}. A natural
question that may arise is how much this correlation can be explain by
user-specifc linguistic patterns. We plan to control for this variable
in future work, for example by building author-conditioned language
models and testing if community vector embeddings can be determined by
the set of the vector embeddings of its members.

>>>>

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurrence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up we aim at testing the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent Bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

@@

All experiments use models with \num{3} sequence encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of parameters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions. 
We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).

@@

\begin{figure*}
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

As a quantitative assessment, we compare the CCLM-learned community
embedding with the community embedding created by \citet{Kumar2018},
who generated them using using a negative-sampling optimization
algorithm and the message-community co-occurrence matrix as ground
truth, using data which spans January 2014 to April 2017.
We refer the reader
to \citet{Kumar2018}
for details, but the important point is that no linguistic
information is used to create these embeddings: they only depend on
whether a given user participates in a given
community. In sharp contrasts, CCLM community embeddings depend in no
way on which user is the author of any given message: we only use the
contents of messages, not authorship data.
%
We correlate the pairwise cosine similarity of the
CCLM embeddings with the same metric on the membership embedding,
restricted to the 46 communities chosen for our study
(\cref{fig:pairwise-comm-sim}).

@@

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Mean community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm-lmcc-ppl}
\end{figure*}

@@

With indiscernibility values evenly ranging over the whole possible
range (0 to 1), we find that CCLMs can be used to recognize certain
communities based on latent Bayesian classification
\ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
distinctive linguistic patterns, none of our model is able to discern
them. In contrast, the distinctive patterns of narrow-topic
communities are just as recognizable by all our LSTM and Transformer
models.

@@

We find that the layer at which the community embedding is taken into
account in a CCLM influences its perplexity \ref{hyp:layer-effect},
albeit weakly. For LSTM models, the perplexity per word, averaged over
all communities, varies between 76.80 and 78.64. (with 80.70 for the unconditioned model)
The perplexity per word is slightly higher for $l_c=0$, and has very
little difference between the other models, with $l_c=1$ being the best.
For transformer models, it varies
between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
function of $l_c$: the maximum is reached at $l_c=2$. We did not
expect non-monotonicity, and leave studying this effect to future work.

The pattern of information gain by community is similar across architectures.
Communities that benefit most from the conditioned model behave that way for 
both the LSTM and transformer.
However, there are some differences.
For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
the best CCLM LSTM but only \num{0.998} for the best transformer model.
Why was the LSTM able to take advantage of the community information for
this community, where the transformer wasn't?
It would be interesting to investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

The representation of communities built by CCLM exhibits a clear
correlation with user co-occurrence patterns
\ref{hyp:extra-linguistic-correlation}. A natural
question that may arise is how much this correlation can be explain by
user-specific linguistic patterns. We plan to control for this variable
in future work, for example by building author-conditioned language
models and testing if community vector embeddings can be determined by
the set of the vector embeddings of its members.
